\bf On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2024-11-23 Manuscript/manuscript.tex
smoothed data section

To compare reaction times at the 2022 World Championships with those
recorded at national competitions, we constructed a dataset of athletes 
who competed in both settings. Initially, we examined data from four 
United States athletes---Trey Cunningham, Daniel Roberts, Grant 
Holloway, and Devon Allen---who participated in both the United States 
Track and Field Championships and the World Championships. Each athlete 
recorded faster reaction times in all World Championships races compared 
to their performances at the national-level event. However, this small 
sample size was insufficient for meaningful statistical analysis. To 
expand the dataset, we included reaction times from additional national 
competitions held between May and July 2022 across various countries. 
Compiling this data was challenging, as reaction times were not centrally 
located and often required searching country-specific websites, with many 
results recorded in native languages.

\jy{Owen, does this include both dash and hurdle? Could you do a
  pooled analysis since there is no gender difference and event
  differece? Need to summarize how many athletes are included in each
  analysis.}

The final dataset consisted of reaction times from athletes who competed 
in both national and international competitions. Reaction times from 
preliminary heats, semifinals, and finals were included to ensure 
sufficient data for analysis. Excluding preliminary heats would have 
significantly reduced the number of athletes and clusters, narrowing the 
scope of the analysis. Each athlete can be considered as a cluster, with
reaction times from the `treatment' group (2022 World Championships) and 
the `control' group (national competition) within the cluster. Cluster 
sizes ranged from three to six, with a median size of four. This
setting forms the basis for rank-based comparison where the
within-cluster dependence needs to be accounted for.

\of{This explanation is going to need to be re-worked now that we have more data.
Should we make a table of some sort? It's going to be very wordy without one}
\eds{We are only doing two analyses with this data, right, one for women 
and one for men?  So perhaps only two sentences needed here? But then 4 sentences
in the next subsection?  Maybe you can just report ranges of cluster sizes and 
median cluster size?}

\subsubsection{2019 and 2023 World Championships}
\label{sec:data2019}
The 2019 and 2023 World Track and Field Championships data is included in the 
larger dataset described in Section~\ref{sec:dataworld}, but we can also use it
similarly to how we used national competitions in Section~\ref{sec:datanational}. 
We can compare the reaction times of athletes who participated in the 2022 World 
Championships with their times at another World Championships: 2019 and/or 2023.

While it is theoretically possible that athletes got faster 
between 2019 and 2022 or 2022 and 2023, it is important to 
remember that these are some of the fastest sprinters on the planet, and thus
improving on elite times is not easy. Once again we repeat a similar procedure
as in Section~\ref{sec:datanational}; we assign each athlete to be a cluster and
within each cluster label times to be from either the 2019, 2022, or 2023 World
Championships. When comparing 2019 and 2022 we look at data from 15
athletes, with two clusters of size two, three clusters of size three, five 
clusters of size four, two clusters of size five, and three clusters of size six.
When comparing 2022 and 2023 we look at data from 22 athletes, with seven 
clusters of size two, three clusters of size three, six clusters of size four,
and six clusters of size five.

\begin{figure}[tbp]
  \centering
  \includegraphics{RankScatterPlots}
  \caption{Reaction times for athletes who competed at the 2022 World Championships
	and at another championship (2022 national, 2019 World, or 2023 World) at which 
	they competed. On the horizontal
  axis below each graph "H", "S", and "F" refer to the heats, semifinals, and 
  finals respectively. Please note that in the last row the 2022 times are to
  the right of the 2023 times.
	\eds{Please check my edited caption; also, I suggest rearranging the rows to 
	use the same ordering as the data was described: put national vs international 
	in first row, 2019 vs 2022 in second row, and 2022 vs 2023 in third row}
	}
  \label{fig:RankScatterplots}
\end{figure}

Figure~\ref{fig:RankScatterplots} shows the data that will be analyzed
with the rank-based methods (described in Section~\ref{sec:rank} with results in 
Section~\ref{subsec:Results_Rank}). In each of the six graphs, the reaction 
times for athletes who competed at the 2022 World Championships
are compared to their reaction times at another at which they competed. 
It is worth noting that for each row of the figure, the athletes compromising 
the 2022 data is different.  This is because the set of athletes who competed at 
both the 2019 and 2022 World Championships is not the same as the set of athletes 
who competed at both the 2022 and 2023 World Championships. %This data was only for athletes who competed
%in at least one race at each championship, however it is possible for athletes
%to have competed in one race in 2019 and two races in 2022. 
Athletes must have competed in at least one race at each championship to be 
included.  The 2022 World
Championship heats and semifinals both had on average lower reaction times than
their respective other (2019, 2023 or 2022 national) races, and the reaction times 
from the 2022 World Championship finals were substantially faster. 
\eds{I would consider removing or rephrasing the last part of the previous 
sentence. With the women's data, it is not as clear the 2022F is lower than 2022S.}
\eds{I also think we need to comment on the men and women's data. But within row
of the figure, don't the distributions appear similar?  If that is so, would a 
reviewer ask why we didn't pool the data for men and women?}
It is also worth noting is that the fastest reaction time in the 2022 World
Championships Finals and Semifinals both came from Devon Allen, but the 
difference between 0.101 and 0.099
seconds was the difference in what caused him to be disqualified.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics{Boxplot}
  \caption{The reaction times from 1999 to 2023 for the men's 110 meter hurdle
  and 100 meter dash.}
  \label{fig:Boxplot}
\end{figure}
\of{Do We want 2 boxplot figures: 1 hurdles, 1 dash or combine them}

>>>>

\jy{Owen, are there names from 100-m dash?}
To compare reaction times at the 2022 World Championships with those
recorded at national competitions, we constructed a dataset of athletes 
who competed in both settings. Initially, we examined data from four
United States athletes---Trey Cunningham, Daniel Roberts, Grant
Holloway, and Devon Allen---who participated in both the United States 
Track and Field Championships and the World Championships. Each athlete 
recorded faster reaction times in all World Championships races compared 
to their performances at the national-level event. However, this small 
sample size was insufficient for meaningful statistical analysis. To 
expand the dataset, we included reaction times from additional national 
competitions held between May and July 2022 across various countries. 
Compiling this data was challenging, as reaction times were not centrally 
located and often required searching country-specific websites, with many 
results recorded in native languages.

The final dataset consisted of reaction times from athletes who competed 
in both national and international competitions. Reaction times from 
preliminary heats, semifinals, and finals were included to ensure 
sufficient data for analysis. Excluding preliminary heats would have 
significantly reduced the number of athletes and clusters. Each
athlete was considered as a cluster, with observed reaction times
from both the `treatment' group (2022 World Championships) and
the `control' group (national competition) within the cluster. Cluster 
sizes ranged from three to six, with a median size of four. Because gender 
is known to influence reaction times \citep{lipps2011implications}, we
prepared data for men and women separately, resulting in 80 reaction
times from 17 athletes for each
gender. This setup forms the basis for the rank-based comparison for
clustered data with subunit-level grouping, which accounts for the
within-athlete dependence inherent in this data structure. The top
panel of Figure!~\ref{fig:RankScatterplots} shows the reaction times
of those who competed both at national competitions and the 2022 World
Championships.

\jy{Could you do a pooled analysis since there is no gender difference
  and event differece in the supplement?}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{RankScatterPlots}
  \caption{Reaction times for athletes who competed at the 2022 World Championships
	and at another championship (2022 national, 2019 World, or 2023 World) at which 
	they competed. On the horizontal
  axis below each graph "H", "S", and "F" refer to the heats, semifinals, and 
  finals respectively. Please note that in the last row the 2022 times are to
  the right of the 2023 times.
	\eds{Please check my edited caption; also, I suggest rearranging the rows to 
	use the same ordering as the data was described: put national vs international 
	in first row, 2019 vs 2022 in second row, and 2022 vs 2023 in third row}
	}
  \label{fig:RankScatterplots}
\end{figure}

\subsubsection{2019 and 2023 World Championships}
\label{sec:data2019}

To compare reaction times at the 2022 World Championships with those 
recorded in the 2019 and 2023 World Championships, we prepared datasets 
including reaction times from athletes who competed in the 2022 World 
Championships and at least one of the 2019 or 2023 World Championships. 
Specifically, we identified athletes who participated in both the 2019 
and 2022 World Championships, as well as those who participated in both 
the 2022 and 2023 World Championships. In the 2019-2022 comparison, 
reaction times from 2022 were treated as the `treatment' group, with 
2019 serving as the `control' group. Similarly, in the 2022-2023 
comparison, reaction times from 2022 were treated as the `treatment' 
group, with 2023 serving as the `control' group. This structure allowed 
us to prepare datasets suitable for examining reaction times of athletes 
who competed across multiple World Championships.

Each athlete was treated as a single cluster, containing their reaction 
times from different World Championships. The dataset for the 2019-2022 
comparison contained 134 reaction times from 34 male athletes and 124 
reaction times from 31 female athletes. The dataset for the 2022-2023 
comparison contained 161 reaction times from 45 male athletes and 182 
reaction times from 47 female athletes. While it is theoretically 
possible that athletes improved their reaction times between 2019 and 
2022 or between 2022 and 2023, such improvements are highly unlikely for 
elite sprinters, as they already operate near the limits of human 
performance. Consequently, consistent improvements observed in 2022 
would suggest systematic differences rather than natural variability.

Figure~\ref{fig:RankScatterplots} shows the reaction times of athletes 
who competed in both the 2019 and 2022 World Championships (middle 
panel) and those who competed in both the 2022 and 2023 World 
Championships (lower panel). The athletes included in the 2022 data 
differ between these two comparisons, as the set of athletes who 
competed in both 2019 and 2022 is not the same as the set who competed 
in both 2022 and 2023. To be included, athletes must have competed in 
at least one race at each championship. Notably, Devon Allen recorded 
the fastest reaction times in both the Finals and Semifinals of the 
2022 World Championships, but his disqualification was determined by a 
difference of just 0.002 seconds, with reaction times of 0.101 and 
0.099 seconds, respectively. This highlights the critical role of 
reaction time precision in elite-level competition.

\eds{I also think we need to comment on the men and women's data. But within row
of the figure, don't the distributions appear similar?  If that is so, would a 
reviewer ask why we didn't pool the data for men and women?}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{Boxplot}
  \caption{The reaction times from 1999 to 2023 for the men's 110 meter hurdle
  and 100 meter dash.}
  \label{fig:Boxplot}
\end{figure}
\of{Do We want 2 boxplot figures: 1 hurdles, 1 dash or combine them}
